---
title: "Practical 2 - Understanding correlation and further exploring relationships"
author: "Team in Room 420 - Megan Ruffle, Naphon Olley, Jennifer James, William Ryan"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

# Intro

---

## Intended Learning Outcomes

After attending this lab, you should be able to use `R` to:

- calculate least squares estimates of model parameters using vector-matrix formulation;
- calculate and interpret the sample correlation coefficient;
- perform hypothesis tests on the population correlation and interpret the decision.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!requireNamespace("webexercises")) {
  stop("You must have the 'webexercises' package installed to knit to HTML.\n   install.packages(\"webexercises\")")
} else {
  library("webexercises")
}
library(webexercises)
library(tinytex)
library(ggplot2)
library(dplyr)
library(knitr)
library(PASWR2)
library(downlit)
library(xml2)
```

```{r eval=FALSE, echo=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

```{r include=FALSE, echo=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

## Introduction

In the lectures we learned how to assess the strength of a linear relationship between random variables using the correlation coefficient.  The population correlation is a measure of the magnitude of the strength of the relationship between two random variables X and Y, and is defined as

\begin{equation}
\rho(X,Y) = \frac{\text{Cov}(X,Y)}{\sqrt{\text{Var}(X)\text{Var}(Y)}},
(\#eq:eq1)
\end{equation}

and can be estimated by replacing each of $\text{Cov}(X,Y)$, \text{Var}(X) and \text{Var}(Y) by their unbiased estimators to give

\begin{equation}
r = \frac{S_{xy}}{\sqrt{S_{xx}S_{yy}}}=\frac{\sum^{n}_{i=1}(x_i-\overline{x})(y_i-\overline{y})}{\sqrt{\sum^{n}_{i=1}(x_i-\overline{x})^2(y_i-\overline{y})^2}},
(\#eq:eq2)
\end{equation}

the **sample correlation coefficient** (-1 $\le r \le$ 1).

Given a sample of data, we can assess the statistical significance of the observed correlations between
variables in the wider population. To do this we perform a hypothesis test (more on this in Chapter \@ref(twopointtwo)).
